---
title: "Using SPRING with R"
output: html_notebook
---

Load necessary pacakges and functions 

```{r}
library(irlba)
library(reticulate)
use_condaenv("anaconda2")

pkl <- import("pickle")
np <- import("numpy")
pd <- import("pandas")

source_python("preprocessing_python.py")
```

Import expression matrix; rows are cells and columns are genes
### ****** Make sure E.npy is unzipped *************

```{r}
print('Loading expression matrix')
E <- np$load('example_inputs/python_E.npy')
dim(E)
```

Filter out cells with fewer than 1000 UMIs

```{r}
print('Filtering cells')
Cell_Filter_Struct <- filter_cells(E,1000)

E <- Cell_Filter_Struct[[1]]
cell_filter <- Cell_Filter_Struct[[2]]
rm(Cell_Filter_Struct)
```

Normalize gene expression data
Only use genes that make up <

```{r}
print('Row-normalizing')
E <- row_normalize(E)
```

Filter genes with mean expression < 0.1 and fano factor < 3

```{r}
print('Filtering genes')
Gene_Filter_Struct <- filter_genes(E,0.1,3)
gene_filter <- Gene_Filter_Struct[[2]]
rm(Gene_Filter_Struct)
```

Z-score the gene-filtered expression matrix and do PCA with 20 pcs

```{r}
print('Zscoring and PCA')
ZS <- Zscore(E[,gene_filter])

# Epca <- get_PCA(ZS, 20)

Epca <- prcomp_irlba(x = ZS, n = 20, retx = TRUE)$x
```

Get euclidean distances in the PC space

```{r}
print('Getting distance matrix')
D <- get_distance_matrix(Epca)
```

load additional data (gene_list, cell_groupings, custom_colors)
gene_list is a list of genes with length E.shape[1]
cell_groupings is a dict of the form: { <grouping_name> : [<cell1_label>, <cell2_label>,...] }
a "grouping" could be the sample id, cluster label, or any other categorical variable
custom_colors is a dict of the form { <color_track_name> : [<cell1_value>, <cell2_value>,...] }
a "custom color" is any continuous variable that you would like to use for coloring cels. 

```{r}
PK <- py_load_object(filename = 'example_inputs/python_data.p', pickle = "pickle")

gene_list <- PK[[1]]
cell_groupings <- PK[[2]]
custom_colors <-  PK[[3]]

names(gene_list) <- 1:length(gene_list)
```

save a SPRING plots with k=5 edges per node in the directory "datasets/frog_python/"
coarse graining can also be performed using the optional coarse_grain_X parameter
     
     
     
```{r}
print('Saving SPRING plot')
save_spring_dir(E,D,5,gene_list,'datasets/frog_python',
                cell_groupings=cell_groupings, custom_colors=custom_colors,
                 coarse_grain_X=1)
```





